\begin{abstract}
In-context learning is a promising approach for offline reinforcement learning (RL) to handle online tasks, which can be achieved by providing task prompts. Recent works demonstrated that in-context RL could emerge with self-improvement in a trial-and-error manner when treating RL tasks as an across-episodic sequential prediction problem. Despite the self-improvement not requiring gradient updates, current works still suffer from high computational costs when the across-episodic sequence increases with task horizons. To this end, we propose an In-context Decision Transformer (IDT) to achieve self-improvement in a high-level trial-and-error manner. Specifically, IDT is inspired by the efficient hierarchical structure of human decision-making and thus reconstructs the sequence to consist of high-level decisions instead of low-level actions that interact with environments. As one high-level decision can guide multi-step low-level actions, IDT naturally avoids excessively long sequences and solves online tasks more efficiently. Experimental results show that IDT achieves state-of-the-art in long-horizon tasks over current in-context RL methods. In particular, the online evaluation time of our IDT is \textbf{36$\bm\times$} times faster than baselines in the D4RL benchmark and \textbf{27$\bm\times$} times faster in the Grid World benchmark.
\end{abstract}

% 尽管不需要额外的梯度计算，但是成本还是很高，局限在short horizon。
% 然后再解释出来为什么成本高，这个过程要明确出low-level action的问题。然后再是我们的方法怎么解决的，要清晰对应上。

% In-context learning is a promising approach for offline reinforcement learning (RL) to handle online tasks, which can be achieved by providing task prompts. Recent works demonstrated that in-context RL could emerge with self-improvement in a trial-and-error manner when treating RL tasks as an across-episodic sequential prediction problem. 
% Despite the advantage of online learning without gradient updates, current works are limited to specific short-horizon tasks due to the high computational cost incurred by across-episodic sequences. Inspired by human decision-making that operates on multiple levels of temporal abstraction, we explore the Hierarchical Chain of experience Context (H2C), which introduces the concept of high-level decisions for in-context RL. H2C learns to improve itself in high-level decisions and then guides the prediction of low-level actions. Unlike existing methods, H2C proposes episodes consisting of high-level decisions, which can avoid excessively long sequences triggered by across-episodic and solve online tasks more efficiently. As we show experiments on the Grid World and D4RL benchmarks, H2C achieves state-of-the-art in almost all tasks, especially outstanding in long-horizon tasks that far exceed recent in-context RL methods. In particular, the online evaluation time of our H2C is 35 times faster than baselines in D4RL benchmarks and 27 times faster in large grid world environments.