We are excited about the opportunities of LLMs as pattern machines for robotics---from reasoning and extrapolating complex patterns as a prior for control, to online optimization of closed-loop policies via sequence improvement. These capabilities present several implications, including (i) perspectives on the role of language pre-training for end-to-end robot learning models \cite{lu2021pretrained,reid2022can}, and (ii) in-context learning of arbitrary patterns as a driving mechanism for policy improvement. LLMs also show promise for mixed autonomy settings---\eg real-time pattern extrapolation for assistive teleoperation. We expect many of these abilities to continue improving as large models expand from learning  patterns within language-only datasets to multimodal domains (\eg images, videos). While this work investigates in-context generalization on fairly simple settings without additional data collection or model training, these capabilities presumably may be significantly improved via domain-specific objectives and finetuning \cite{li2022pre,driess2023palm,zhang2023motiongpt,lee2023supervised, dinh2022lift}.


\textbf{Limitations \& Future Work.} Today, the inference costs (and monetary costs) of using LLMs in the control loop are quite high. Predicting the next token for every sequence, \eg every dimension of every time step in a trajectory, involves querying an LLM. State-action spaces which are higher dimensional and/or greater precision also result in longer  representations, and thereby the extent to which they can be extrapolated or sequence optimized is bounded by the context length of models. These limitations may prevent deploying these models on more complex tasks in practice; however, they may be partially mitigated by incorporating mechanisms like external memory, and by current efforts to drive improvements in LLM quantization \cite{zafrir2019q8bert} and inference efficiency \cite{dao2022flashattention}. An additional limitation lies in the fact that, for best performance, some care must be taken to represent patterns with consistent tokenization (which requires knowledge of the model's tokenization scheme). Finally, as with any other language-only model, LLM-based control may (i) be unpredictable, and (ii) lack visual/physical grounding; thus, it is not currently suitable for application outside of constrained lab settings. We leave the exploration of these important topics for future work.
